---
title: "Crimes_Vancouver"
author: "Andy Liang"
date: "08/06/2021"
output: html_document
---

#last edited
```{r}
#2020.1.14
```


```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
crime_data = read_csv("crimedata_csv_all_years.csv")
```

#In this code chunk, I removed observations with missing data, created a new dataframe with a date column, changed neighbourhood and type of crime into categorical data. Because I'm only interested in the year 2020 the data is subset to only 2020.
```{r}
crime_data=na.omit(crime_data)
summary(crime_data)
str(crime_data)
crime_dates = crime_data %>% dplyr::select(TYPE:MINUTE,NEIGHBOURHOOD) %>% mutate(TOTALMINS = HOUR*60 + MINUTE) %>% mutate(DATE = make_datetime(YEAR,MONTH,DAY,HOUR,MINUTE))
crime_dates$TYPE = as.factor(crime_dates$TYPE)
crime_dates$YEAR = as.factor(crime_dates$YEAR)
crime_dates$NEIGHBOURHOOD = as.factor(crime_dates$NEIGHBOURHOOD)
crime_df = crime_dates %>% filter(YEAR == 2020)
```




#Making a new column for the most frequent type of crime by time of the day
```{r}
#[Inclusive Start,Exclusive end)
INTERVALS = c("12am-3am","3am-6am","6am-9am","9am-12pm","12pm-3pm","3pm-6pm","6pm-9pm","9pm-12am")
total_mins = c(0,179,359,539,719,899,1079,1259,1440)
#created new column for interval
data_intervals = data.frame(INTERVALS[findInterval(crime_df$TOTALMINS,total_mins)])
#rename the column
colnames(data_intervals)="T_PERIOD"
```


#column bind the time period with the dataset
```{r}
crime_df = cbind(crime_df,data_intervals)
crime_df$T_PERIOD = as.factor(crime_df$T_PERIOD)
crime_df
```

#Data Visualization

```{r}
crime_df %>% ggplot(aes(NEIGHBOURHOOD)) + geom_bar(aes(fill=TYPE)) + coord_flip()

#looking at this barplot we can see that the Central Business District is the neighborhood with the most type of crimes
```







#Preparation for the Heat map of type of crime vs time of crime
```{r}
crime_df2 = crime_df %>% dplyr::select(TYPE,T_PERIOD) %>% group_by(T_PERIOD) %>% mutate(T_PERIOD = factor(T_PERIOD,levels=INTERVALS)) %>% arrange(T_PERIOD) %>% count(TYPE)
crime_df2
library(reshape)
hmapdata = as.matrix(cast(crime_df2,T_PERIOD~TYPE,value="n"))
hmapdata[is.na(hmapdata)]=0
colrs=colorRampPalette(c("white","maroon"))
heatmap(hmapdata,Rowv = NA,Colv = NA,col=colrs(300),ylab="Time",xlab="Type of Crime",cexRow = 0.65,cexCol = 0.65,margins = c(15,7),main="Types of Crime vs Time of Crime ")
```




#Heat Map of Vancouver's Type of Crime  vs Neighborhood 
```{r}
crime_df3 = crime_df %>% dplyr::select(TYPE,NEIGHBOURHOOD) %>% group_by(NEIGHBOURHOOD) %>% count(TYPE) 
hmapdata2 = as.matrix(cast(crime_df3,TYPE~NEIGHBOURHOOD,value="n"))
hmapdata2[is.na(hmapdata2)]=0
colrs2=colorRampPalette(c("white","maroon"))
heatmap(hmapdata2,Rowv = NA,Colv = NA,col=colrs2(300),xlab="Neighborhood",ylab="Type of Crime",cexRow = 0.65,cexCol = 0.65,margins = c(10,15),main="Types of Crime vs Neighborhood ") 
```


#creating a map


#used to convert utm into lat and long vals.
```{r}
library(rgdal)
```

```{r}
locations = as.data.frame(cbind(crime_data$X,crime_data$Y))
sputm <- SpatialPoints(locations, proj4string=CRS("+proj=utm +zone=10 +datum=WGS84")) 
spgeo <- spTransform(sputm, CRS("+proj=longlat +datum=WGS84"))
mypoints=data.frame(spgeo@coords)
```

#The crime dataset with longitude and latitude data.
```{r}
crime.dataset = cbind(crime_data,mypoints)
```


```{r}
library(data.table)
library(leaflet)
library(sp)
library(KernSmooth)
library(rgdal)
library(MASS)
library(raster)
```


```{r}
crimedata2020 = crime.dataset %>% filter(YEAR==2020)
setnames(crimedata2020, tolower(colnames(crimedata2020)))
colnames(crimedata2020)[11] = "longitude"
colnames(crimedata2020)[12] ="latitude"
head(crimedata2020)

```

#removing garbage datapoints
```{r}
crimedata2020 = crimedata2020 %>% filter(longitude !=0)
crimedata2020 = crimedata2020 %>% filter(latitude !=0)

```


```{r}
MASS::bandwidth.nrd(as.vector(crimedata2020$longitude)) #0.01561392
MASS::bandwidth.nrd(as.vector(crimedata2020$latitude))# 0.01025273
```


```{r}
crimedata2020 = as.data.table(crimedata2020)
```

#taking a sample to make a plot as 33,000 points is too much obserations for a map

```{r}
mysample = crimedata2020[sample(nrow(crimedata2020), 3000), ]
```


#contour lines
```{r}

the_kde <- bkde2D(crimedata2020[ ,list(longitude,latitude)],
              bandwidth=c(0.0018, 0.00115247), gridsize = c(300,300))

kernel_d_raster <- raster(list(x=the_kde$x1 ,y=the_kde$x2 ,z = the_kde$fhat))


kernel_d_raster@data@values[which(kernel_d_raster@data@values < 15)] <- NA


palRaster <- colorBin("Spectral",bins=10, domain = kernel_d_raster@data@values, na.color = "transparent")

## Leaflet map with raster
leaflet() %>% addTiles() %>% 
  addRasterImage(kernel_d_raster, 
                 colors = palRaster, 
                 opacity = .4) %>%
  addLegend(pal = palRaster, 
            values = kernel_d_raster@data@values, 
            title = "Kernel Density of Points")
```


