---
title: "Crimes_Vancouver"
author: "Andy Liang"
date: "08/06/2021"
output: html_document
---

#Purpose
```{r}
#Most frequent time for different crimes in vancouver throughout the years.

```


```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)

crime_data = read_csv("crimedata_csv_all_years.csv")
```

```{r}
crime_data=na.omit(crime_data)
summary(crime_data)
str(crime_data)


crime_dates = crime_data %>% select(TYPE:MINUTE,NEIGHBOURHOOD) %>% mutate(TOTALMINS = HOUR*60 + MINUTE) %>% mutate(DATE = make_datetime(YEAR,MONTH,DAY,HOUR,MINUTE))

crime_dates$TYPE = as.factor(crime_dates$TYPE)
crime_dates$YEAR = as.factor(crime_dates$YEAR)
crime_dates$NEIGHBOURHOOD = as.factor(crime_dates$NEIGHBOURHOOD)
crime_dates


```




#Most Frequent Type of Crime by time of the day by year
```{r}
#[Inclusive Start,Exclusive end)
INTERVALS = c("12am-3am","3am-6am","6am-9am","9am-12pm","12pm-3pm","3pm-6pm","6pm-9pm","9pm-12am")

total_mins = c(0,179,359,539,719,899,1079,1259,1440)

#created new column for interval
data_intervals = data.frame(INTERVALS[findInterval(crime_dates$TOTALMINS,total_mins)])

#rename the column
data_intervals=rename(data_intervals,`T_PERIOD`=`INTERVALS.findInterval.crime_dates.TOTALMINS..total_mins..`)

```


#New df for stuff i need.
```{r}
crime_df = cbind(crime_dates,data_intervals)
crime_df$T_PERIOD = as.factor(crime_df$T_PERIOD)

crime_df


```

#Preparation for the Heat map of type of crime vs time of crime
```{r}

crime_df2 = crime_df %>% select(TYPE,T_PERIOD) %>% group_by(T_PERIOD) %>% mutate(T_PERIOD = factor(T_PERIOD,levels=INTERVALS)) %>% arrange(T_PERIOD) %>% count(TYPE)

crime_df2




```

#Heat Map of Vancouver's Type of Crime  vs Time of Crime
```{r}
library(reshape)
hmapdata = as.matrix(cast(crime_df2,T_PERIOD~TYPE,value="n"))

hmapdata[is.na(hmapdata)]=0
colrs=colorRampPalette(c("white","maroon"))
heatmap(hmapdata,Rowv = NA,Colv = NA,col=colrs(300),ylab="Time",xlab="Type of Crime",cexRow = 0.65,cexCol = 0.65,margins = c(15,7),main="Types of Crime vs Time of Crime ")


```


#Preparataion for the Heat map of Type of crime vs Neighborhood
```{r}

crime_df3 = crime_df %>% select(TYPE,NEIGHBOURHOOD) %>% group_by(NEIGHBOURHOOD) %>% count(TYPE)

crime_df3

```



#Heat Map of Vancouver's Type of Crime  vs Neighborhood 
```{r}
hmapdata2 = as.matrix(cast(crime_df3,TYPE~NEIGHBOURHOOD,value="n"))

hmapdata2[is.na(hmapdata2)]=0
colrs2=colorRampPalette(c("white","maroon"))
heatmap(hmapdata2,Rowv = NA,Colv = NA,col=colrs2(300),xlab="Neighborhood",ylab="Type of Crime",cexRow = 0.65,cexCol = 0.65,margins = c(10,15),main="Types of Crime vs Neighborhood ")

```




#used to convert utm into lat and long vals.
```{r}
library(rgdal)
```

```{r}

locations = as.data.frame(cbind(crime_data$X,crime_data$Y))

sputm <- SpatialPoints(locations, proj4string=CRS("+proj=utm +zone=10 +datum=WGS84")) 
spgeo <- spTransform(sputm, CRS("+proj=longlat +datum=WGS84"))

mypoints=data.frame(spgeo@coords)
```

#The crime dataset with longitude and latitude data.
```{r}
crime.dataset = cbind(crime_data,mypoints)

```


#filtering to 2020 and 2019
```{r}
head(crime.dataset)

crimedata2019Q1 = crime.dataset %>% filter(YEAR==2019 & MONTH ==1) 


crimedata2020 = crime.dataset %>% filter(YEAR==2020)

```



#libraries for a map
```{r}
library(sf)
library(mapview)
```


```{r}
table(crimedata2019Q1$NEIGHBOURHOOD)
```



#MAP of Q1 data
```{r}

mapdata2019Q1 = data.frame(cbind(crimedata2019Q1$TYPE,crimedata2019Q1$V1,crimedata2019Q1$V2))
mapdata2019Q1$X1 = as.factor(mapdata2019Q1$X1)
mapdata2019Q1$X2 = as.numeric(mapdata2019Q1$X2)
mapdata2019Q1$X3 = as.numeric(mapdata2019Q1$X3)
mapdata2019Q1


locations.sf = st_as_sf(mapdata2019Q1,coords=c("X2","X3"),crs=4326)


mapview(locations.sf,cex="X1",alpha=0.2,legend=TRUE)


```


#MERGED DATA
```{r}
merged_data = cbind(crime_df,mypoints)

merged_data = merged_data %>% select(TYPE,YEAR,MONTH,TOTALMINS,T_PERIOD,V1,V2) %>% filter(YEAR ==2019 | YEAR == 2020)


table(merged_data$TYPE)

head(merged_data)
```


#MAP OF Q1 data about type of crime during different parts of a day

```{r}
map_2019_Q1_0to3am = merged_data %>% select(TYPE,YEAR,MONTH,T_PERIOD,V1,V2) %>% filter(YEAR==2019 & MONTH <=6 & T_PERIOD == "12am-3am" ) 

map_2020_Q1_0to3am = merged_data %>% select(TYPE,YEAR,MONTH,T_PERIOD,V1,V2) %>% filter(YEAR==2020 & MONTH <=6 & T_PERIOD == "12am-3am" ) 
```



```{r}
map_2019_Q1_0to3am. = map_2019_Q1_0to3am[,c(1,5,6)] 



locations.sf = st_as_sf(map_2019_Q1_0to3am.,coords=c("V1","V2"),crs=4326)


mapview(locations.sf,cex="TYPE",legend=TRUE, map.types="OpenStreetMap")



```


```{r}
map_2020_Q1_0to3am. = map_2020_Q1_0to3am[,c(1,5,6)] 



locations.sf = st_as_sf(map_2020_Q1_0to3am.,coords=c("V1","V2"),crs=4326)


mapview(locations.sf,cex="TYPE",legend=TRUE,map.types="OpenStreetMap")


```



```{r}
library(leaflet)
```


```{r}
head(map_2019_Q1_0to3am.)
map_2019_Q1_0to3am.$V1 = as.numeric(map_2019_Q1_0to3am.$V1)
map_2019_Q1_0to3am.$V2 = as.numeric(map_2019_Q1_0to3am.$V2)

mypoints2=data.frame(cbind(map_2019_Q1_0to3am.$V1,map_2019_Q1_0to3am.$V2))



sp.data = SpatialPointsDataFrame(mypoints2,map_2019_Q1_0to3am.)

map = leaflet() %>% addTiles() %>% addMarkers(data = sp.data,lng = ~V1,lat=~V2,popup = ~TYPE)

map
```
